---
title: "Lab 02"
subtitle: "CS3172-1, Spring 2023, Effat University"
author: "Aicha Sidiya"
editor_options: 
  chunk_output_type: inline
output: pdf_document
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
library(scales)
``` 

# Data

```{r read-in-data, message = FALSE}
cas <- read_rds("data/canada_survey.rds")
```

# Tasks

## Task 1

```{r}
cas <- filter(cas, energy_expense > 0, household_income > 0)
```

```{r}
cas <- mutate(cas, marital_status = factor(marital_status))
```

## Task 2

```{r}
cas <-
  mutate(cas, heat_equip = case_when(
    heat_equip == 1 ~ "steam",
    heat_equip == 2 ~ "forced air",
    heat_equip == 3 ~ "stove",
    heat_equip == 4 ~ "electric heating"
  ))

```

```{r}
cas <-
  mutate(cas, heat_fuel = case_when(
    heat_fuel == 1 ~ "oil",
    heat_fuel == 2 ~ "gas",
    heat_fuel == 3 ~ "electricity",
    heat_fuel == 4 ~ "other"
  ))

```

## Task 3

```{r}
cas %>% 
  group_by(heat_fuel, heat_equip) %>% 
  summarize(mean_expense = mean(energy_expense), 
            median_expense = median(energy_expense),
            stdv_expense = sd(energy_expense))
```

* **What combination of fuel type and equipment has the highest average energy expense?**

Oil and Forced air have the highest mean which is 3498.850. 

* **Which combination has the most variability with regards to energy expense?**

Other	and Steam have the highest variability with a standard deviation 2278.911.

*	**Which type of heating equipment doesn’t take all possible fuel types?**

The electric heating equipment use only electricity as fuel. 


## Task 4

```{r}
ggplot(data=cas, 
       mapping=aes(x=heat_fuel, y=energy_expense)) + 
  geom_boxplot() +
  facet_wrap(.~heat_equip, nrow = 2)+
  scale_y_continuous(labels = scales :: dollar_format())+
  labs(title = "Energy Expense by Heating Fuel Type",
       subtitle = "faceted by type of heating equipment", 
       x="Heat Fuel Type", 
       y="Energy Expenses") 
  
```

## Task 5

```{r}
cas <- mutate(cas, prop_energy_expense = 
                (energy_expense / household_income))
cas2 <- cas %>% 
  mutate(prop_energy_expense = 
           (energy_expense / household_income)) %>%
  arrange(desc(prop_energy_expense)) %>%
  slice(1,n()) %>% glimpse()
#slice(1,n()) getting first and last column
```
The respondent with the lowest proportion of energy expense per household income which is 0.0000149 is a 85 year old divorced or widowed female that has a household income of 67000 $ and no degree, she lives in an apartment in Ontario that was built in 1971-1980, with 6 rooms, 2 beds and 1 bath she uses forced air heating equipment.  

The respondent with the highest proportion of energy expense per household income 37.8 is a 55 year old divorced female that has a household income of 100 $ wih a bachelor of university diplome , she lives in a single detached in Saskatchewan that was built in 1971-1980, with 7 rooms, 3 beds and 1 bath she uses forced air heating equipment. 

## Task 6

```{r message=FALSE}
cas %>% 
  group_by(year, province) %>% 
  summarise(median_energy_per_room = 
              median(energy_expense/rooms)) %>%
  slice_min(median_energy_per_room)
```

## Task 7

```{r fig.width=7, fig.height=6, message=FALSE}
cas %>% mutate(energy_vulnerable = 
                ifelse(prop_energy_expense > 0.05, "Yes", "No")) %>%
  group_by(education, age) %>%
  summarise(prop_vulnerable = 
              mean(energy_vulnerable == "Yes")) %>%
  ungroup() %>%
ggplot(aes(x = education, 
           y = age, 
           fill = prop_vulnerable)) +
  geom_raster() +
  scale_fill_gradient( low = "white", 
                       high = "red") +
  theme_bw() +
  labs(x = "Education", 
       y = "Age", 
       fill = "Proportion energy vulnerable")
```

## Task 8

